---
title: "Overlap Analysis"
output:
  html_document:
    df_print: paged
---

# WJS implementation

```{r}
fill_set <- function(a, b) {
    names_a_minus_b <- setdiff(names(b), names(a))
    la <- length(a)
    la_miss <- length(names_a_minus_b)
    r_miss <- ((la + la_miss)*(la + la_miss + 1) - la*(la+1))/(2*la_miss)
    miss <- rep(r_miss, la_miss)
    names(miss) <- names_a_minus_b
    #return
    c(a, miss)
}

WJS <- function(a, b) {
    a <- fill_set(a, b)
    b <- fill_set(b, a)
    b <- b[names(a)]
    d <- cbind(a, b)
    #su colonne
    d <- apply(d, 2, function(x) length(x) - x + 1)
    #su righe
    res <- apply(d, 1, function(x) c(min(x), max(x)))
    #return
    sum(res[1, ])/sum(res[2, ])
}
```

# Option 2

## Procedure

Map isoforms to genes last:

1. map genes associated to CAD (Opentarget) to their isoforms. Rank of all
isoforms of a gene = rank of the gene
2. use weighted version of Jaccard index to measure overlap between the each
backward expansion (in isoforms) and list of CAD isoforms
3. use permutational approach to issue p-values
4. find isoforms with a significant overlap, then map them to genes

## Implementation

### Local files

```{r}
cad_genes_file <-"Open_Target.csv"
iso_mapping_file <- "Iso-Gene_Mapping.csv"
nessra_dir <- "./Already_In_Boinc/"
```

### Preprocess data

```{r}
cad_genes <- read.csv(cad_genes_file, sep = ',')[, c(1, 4)]
isoforms_map <- read.csv(iso_mapping_file, sep = '\t', stringsAsFactors = FALSE)
colnames(cad_genes) <- c('gene', 'association_score')

all_isoforms <- unique(isoforms_map$isoform)

# simpler way to map cad gene to isoforms

cad_isoform_mapping <- merge(cad_genes, isoforms_map, by = 'gene')
ranked_cad_isoforms <- rank(cad_isoform_mapping$association_score)
ranked_cad_isoforms <- length(ranked_cad_isoforms) - ranked_cad_isoforms + 1
names(ranked_cad_isoforms) <- cad_isoform_mapping$isoform

# isoform to gene mapping, useful for mapping significant isoforms to genes

isoform_gene_mapping <- isoforms_map$gene
names(isoform_gene_mapping) <- isoforms_map$isoform

# read nessra expansion files

expansion_files <- paste0(nessra_dir, list.files(nessra_dir))
nessra_iso_results <- lapply(expansion_files, read.csv, skip = 1,
    stringsAsFactors = FALSE)

# function to format data from nessra expansion files

library(stringr)

get_expansion_form_file <- function(file) {
  
  iso_names <- str_replace(file$node, "t", "T")
  iso_ranks <- rank(file$Frel)
  names(iso_ranks) <- iso_names
  length(iso_ranks) - iso_ranks + 1
}

nessra_iso_results <- lapply(nessra_iso_results, get_expansion_form_file)

# name each ranked expansion vector with the name of the expanded isoform

isoform_names <- lapply(expansion_files, readLines, n = 1)
isoform_names <- str_match(isoform_names, "T[0-9]{6}")
names(nessra_iso_results) <- isoform_names
```

## Compute WJS scores

```{r}
WJS_results <- vapply(nessra_iso_results, WJS, numeric(1), ranked_cad_isoforms)
```


## Create null distributions
```{r}
# len = length of the nessra expansion i.e. backward expansion
# times = how many random samples
# ref = reference gene list, i.e. CAD isoforms i.e.
# set = set of names to sample from, i.e. names in the FANTOM dataset
all_iso<-unique(isoforms_map$isoform)

make_null_cdf <- function(len, times, ref, set) {
    random_samples <- replicate(times, {
        random_sample <- seq_len(len)
        names(random_sample) <- sample(set, len)
        random_sample
    }, simplify = FALSE)
    
    ecdf(vapply(random_samples, WJS, numeric(1), ref))
}
```

### Computing null distributions for each length.

```{r}
library(parallel)

lens <- vapply(nessra_iso_results, length, numeric(1))

times <- 2000

cores <- 7 # number of cores

# Go parallel
cl <- makePSOCKcluster(cores)
clusterExport(cl, c("WJS", "fill_set")) # make these visible to workers
null_distributions <- parLapply(cl, unique(lens), make_null_cdf,
    times, ranked_cad_isoforms, all_isoforms)
stopCluster(cl)

names(null_distributions) <- unique(lens)
```

### Computing p_values.

```{r}
# you only extract the distribution for a certain length
p_values <- Map(function(x, y) 1 - null_distributions[[as.character(y)]](x),
    WJS_results, lens)
# multiple hypothesis test correction
p_adjusted <- p.adjust(p_values, "BH")

# significant genes, alpha to be decided
alpha <- 0.05
significant_isoforms <- names(WJS_results)[p_adjusted < alpha]
significant_genes <- isoform_gene_mapping[significant_isoforms]
```

##Save data

```{r}
save(WJS_results, p_values, p_adjusted, null_distributions,
    file = "Option2_already_in_boinc.RData")
```
