---
title: "Overlap Analysis"
output: html_notebook
---

# WJS implementation

```{r}
fill_set <- function(a, b) {
    names_a_minus_b <- setdiff(names(b), names(a))
    la <- length(a)
    la_miss <- length(names_a_minus_b)
    r_miss <- ((la + la_miss)*(la + la_miss + 1) - la*(la+1))/(2*la_miss)
    miss <- rep(r_miss, la_miss)
    names(miss) <- names_a_minus_b
    #return
    c(a, miss)
}

WJS <- function(a, b) {
    a <- fill_set(a, b)
    b <- fill_set(b, a)
    b <- b[names(a)]
    d <- cbind(a, b)
    #su colonne
    d <- apply(d, 2, function(x) length(x) - x + 1)
    #su righe
    res <- apply(d, 1, function(x) c(min(x), max(x)))
    #return
    sum(res[1, ])/sum(res[2, ])
}
```

# Option 2

## Procedure

Map isoforms to genes last:

1. map genes associated to CAD (Opentarget) to their isoforms. Rank of all isoforms of a gene
= rank of the gene. (?) Do scaling of the ranks   
#### no information on the most significant isoforms---> taking more isoforms for the first genes? no use maybe.
2. use weighted version of Jaccard index to measure overlap between the each backward expansion (in isoforms) and
list of CAD isoforms
3. use permutational approach to issue p-values
4. find isoforms with a significant overlap, then map them to genes

## Implementation

```{r}
cad_genes<- read.csv('targets_associated_with_coronary_artery_disease.csv',header = TRUE,sep=',')[,c(1,3)]
isoforms_map <-read.csv('Iso-Gene_Mapping.csv',header = TRUE,sep='\t')
colnames(cad_genes) <- c('gene','overall_score')


n<-data.frame(gene = rep(NA,48857), overall_score = rep(NA,48857))
rownames(n)<-c(871:49727)
cad_genes<-rbind(cad_genes,n)
map<-merge(cad_genes,isoforms_map,by='gene')
map<-map[is.na(map$overall_score)==FALSE,]
#naming genes with unique isoforms
named_map<-map[1:2]
rownames(named_map)<-map$isoform
#now I have my mapped dataframe 

sorted_map<-named_map[order(named_map$overall_score,decreasing=T),]
#now I have my ordered mapped dataframe 

#compute ranks for the WJS
ranked<-rank(sorted_map$overall_score)
names(ranked)<-rownames(sorted_map)
r_ranked<- length(ranked) - ranked + 1


```


```{r}
nessra_dir <- "./subset/"
expansion_files <- paste0(nessra_dir, list.files(nessra_dir))
nessra_gene_results <- lapply(expansion_files, read.csv, skip = 1,  #list of files to apply WJS to
    stringsAsFactors = FALSE)

```
## "fromExpansion" function to interact with files, "fromWJS" to collect scores of Jaccard similarity
```{r}
library(stringr)
fromExpansion<-function(file){
  
  isos<-str_replace(file$node,"t","T")
  rk<-file$rank
  names(rk)<-isos
  rk 
}

toWJS<-lapply(nessra_gene_results,fromExpansion)

fromWJS<-vapply(toWJS,WJS,numeric(1),r_ranked)
#now I have my vector of WJS scores

isoform_names <- lapply(expansion_files, readLines, n = 1)
isoform_names <- str_match(isoform_names, "T[0-9]{6}") 
#associa ad ogni ranking list il nome dell'isoforma 
names(fromWJS) <- isoform_names

```

## Create null distributions
```{r}
# len = length of the nessra expansion i.e. backward expansion
# times = how many random samples
# ref = reference gene list, i.e. CAD isoforms i.e.
# set = set of names to sample from, i.e. names in the FANTOM dataset
all_iso<-unique(isoforms_map$isoform)

make_null_cdf <- function(len, times, ref, set) {
    random_samples <- replicate(times, {
        random_sample <- seq_len(len)
        names(random_sample) <- sample(set, len)
        random_sample
    }, simplify = FALSE)
    
    ecdf(vapply(random_samples, WJS, numeric(1), ref))
}
```

### Computing null distributions for each length.

```{r}
library(parallel)

lens <- vapply(toWJS, length, numeric(1))

times <- 2000

cores <- 7 # number of cores

# Go parallel
cl <- makePSOCKcluster(cores)
clusterExport(cl, c("WJS", "fill_set")) # make these visible to workers
null_distributions <- parLapply(cl, unique(lens), make_null_cdf,
    times, r_ranked, all_iso)
stopCluster(cl)

names(null_distributions) <- unique(lens)
```

### Computing p_values.

```{r}
#you only extract the distribution for a certain len already computed
p_values <- Map(function(x, y) 1 - null_distributions[[as.character(y)]](x),
    fromWJS, lens)
# multiple hypothesis test correction
p_adjusted <- p.adjust(p_values, "BH")

# significant genes, alpha to be decided

p_vec2<-c(names(p_adjusted[p_adjusted<0.05]))
ind2<-match(p_vec,isoforms_map[,1])
res_genes2<-unique(as.character(isoforms_map[ind2, 2]))

```

##Save data

```{r}

save(res_genes,fromWJS, p_values, p_adjusted, null_distributions,
    file = "Option2.RData")
```
